{\scshape Denicek}: Computational Substrate for\\ Document-Oriented End-User Programming

### 2025-07-19 paper.tex
Add ACM submission

\begin{document}
\title[Denicek: Computational Substrate for Document-Oriented End-User
  Programming]{{\scshape Denicek}: Computational Substrate for\\ Document-Oriented End-User Programming\\[0.25em]}

@@

\keywords{Programming Systems, Computational Substrate, End-User Programming,
 Programming by Demonstration, Local-First Software}

\begin{teaserfigure}
  \vspace{-0.2em}
  \includegraphics[width=1\textwidth,clip,trim=0cm 8.5cm 7.7cm 0cm]{fig/teaser.pdf}
  \vspace{-1.5em}
  \caption{\diff{Denicek is a computational substrate for document-oriented programming
    based on document edit histories~(A). We co-design Denicek with a web-based end-user
    programming environment Denicek (B). Here, Denicek is used to build a Todo list app via
    programming-by-demonstration, by copying a value from input to a new list item.
    We then evaluate the generality of Denicek by using it to build a data science notebook
    system Datnicek (C). Here, Datnicek is used to interactively clean the data table, by
    removing the `` p'' (provisional) marker from numerical columns.
  }}
  \Description{The Denicek diagram in the middle illustrates a document representation as an
   edit history. Two arrows point from the diagram to two screenshots. Webnicek on the left
   contains a Todo list app with a command toolbox showing add item operation. Datnicek on the
   right contains two cells, one with code to load data and another with an interactive editor
   for data cleaning.}
  \label{fig:teaser}
  \vspace{0.75em}
\end{teaserfigure}
\maketitle

@@

\begin{figure}
\newcommand{\extablecol}[5]{
\small{\bfseries #1} #2\; --\; #3\\
\quad \footnotesize #4 \\[-0.2em]
\quad \footnotesize \emph{#5} \\[0.6em]
}
\hspace*{-0.32cm}\begin{tabular}{|l|}
\hline
\\[-1em]
\extablecol{Counter App}{\cite{kiss-2014-7guis}}
  {Counter with increment and decrement buttons.}
  {The current count is represented by a formula that is modified by the buttons.\\[-0.2em]\quad\footnotesize
  The user can inspect the evaluation trace to see how the count was modified.}
  {Programming by Demonstration, Incremental Recomputation, End-User Debugging}
\extablecol{Todo App}{\cite{osmani-2024-todomvc}}
  {Buttons to add an item and remove all completed.}
  {Adding an item must correctly merge with independently added functionality to \\[-0.2em]\quad\footnotesize
   compute which items are completed and remove them based on a formula result.}
  {Programming by Demonstration, Collaborative Editing, Incremental Recomputation}
\extablecol{Conference List}{\cite{edwards-2025-schema}}
  {Manage a list of invited conference speakers.}
  {Adding speakers to a list through an in-document user interface merges with \\[-0.2em]\quad\footnotesize
   refactoring that turns the list into a table and separates name from an email.}
  {Local-First Collaboration, Programming by Demonstration}
\extablecol{Conference Budget}{\cite{edwards-2025-schema}}
  {Calculate budget based on a speaker list.}
  {References are updated  when the list is refactored. Only affected formulas\\[-0.2em]\quad\footnotesize
   are recomputed and the user can view elements on which the result depends.}
  {Local-First Collaboration, Incremental Recomputation, End-User Debugging}
\extablecol{Hello World}{\cite{miller-2001-simult}}
  {Normalize the capitalization of two word messages.}
  {An operation to normalize the text in a list item can be recorded and applied to\\[-0.2em]\quad\footnotesize
   all list items or, alternatively, applied directly to all list items.}
  {Programming by Demonstration}
\extablecol{Traffic Accidents}{\cite{edwards-2022-copypaste}}
  {Compute statistics using two data sources.}
  {Formula to compute statistics can be reused with a different data source;\\[-0.2em]\quad\footnotesize
   error correction is propagated automatically to the copied version of the formula.}
  {Concrete Programming, Incremental Recomputation}
\hline
\end{tabular}

\caption{Formative examples used in Denicek design}
\label{fig:examples}
\end{figure}

>>>>

\begin{document}
\title[Denicek: Computational Substrate for Document-Oriented End-User
  Programming]{{\scshape Denicek}: Computational Substrate for\\ Document-Oriented End-User Programming}

@@

\begin{CCSXML}
 <ccs2012>
    <concept>
        <concept_id>10011007.10011006.10011066</concept_id>
        <concept_desc>Software and its engineering~Development frameworks and environments</concept_desc>
        <concept_significance>500</concept_significance>
        </concept>
    <concept>
        <concept_id>10003120.10003121.10003129.10011756</concept_id>
        <concept_desc>Human-centered computing~User interface programming</concept_desc>
        <concept_significance>300</concept_significance>
        </concept>
    <concept>
        <concept_id>10003120.10003121.10003124.10010868</concept_id>
        <concept_desc>Human-centered computing~Web-based interaction</concept_desc>
        <concept_significance>300</concept_significance>
        </concept>
  </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~Development frameworks and environments}
\ccsdesc[300]{Human-centered computing~User interface programming}
\ccsdesc[300]{Human-centered computing~Web-based interaction}

\keywords{Programming Systems, End-User Programming}
%Computational Substrate,  Programming by Demonstration, Local-First Software}

\begin{teaserfigure}
  \vspace{0.25em}
  \includegraphics[width=1\textwidth,clip,trim=0cm 8.5cm 7.7cm 0cm]{fig/teaser.pdf}
  \vspace{-1.5em}
  \caption{\diff{Denicek is a computational substrate for document-oriented programming
    based on document edit histories~(A). We co-design Denicek with a web-based end-user
    programming environment Denicek (B). Here, Denicek is used to build a Todo list app via
    programming-by-demonstration, by copying a value from input to a new list item.
    We then evaluate the generality of Denicek by using it to build a data science notebook
    system Datnicek (C). Here, Datnicek is used to interactively clean the data table, by
    removing the `` p'' (provisional) marker from numerical columns.
  }}
  \Description{The Denicek diagram in the middle illustrates a document representation as an
   edit history. Two arrows point from the diagram to two screenshots. Webnicek on the left
   contains a Todo list app with a command toolbox showing add item operation. Datnicek on the
   right contains two cells, one with code to load data and another with an interactive editor
   for data cleaning.}
  \label{fig:teaser}
  \vspace{0.75em}
\end{teaserfigure}
\maketitle

@@

\begin{figure*}
\newcommand{\extablecol}[5]{
\small{\bfseries #1} #2\; --\; #3\\
\quad \footnotesize #4 \\[-0.2em]
\quad \footnotesize Programming experiences: \emph{#5} \\[0.6em]
}
\begin{tabular}{|l|}
\hline
\\[-1em]
\extablecol{Counter App}{\cite{kiss-2014-7guis}}
  {Counter with increment and decrement buttons.}
  {The current count is represented by a formula that is modified by the buttons.
  The user can inspect the evaluation trace to see how the count was modified.}
  {Programming by Demonstration, Incremental Recomputation, End-User Debugging}
\extablecol{Todo App}{\cite{osmani-2024-todomvc}}
  {Todo with buttons to add an item and remove all completed items.}
  {Adding an item must correctly merge with independently added functionality to
   compute which items are completed and remove them based on a formula result.}
  {Collaborative Editing, Programming by Demonstration, Incremental Recomputation, Schema Change Control}
\extablecol{Conference List}{\cite{edwards-2025-schema}}
  {Managing a list of invited conference speakers and schema change.}
  {Adding speakers to a list through an in-document user interface merges with
   refactoring that turns the list into a table and separates name from an email.}
  {Collaborative Editing, Programming by Demonstration, Schema Change Control}
\extablecol{Conference Budget}{\cite{edwards-2025-schema}}
  {Calculate budget based on a speaker list or a table.}
  {References are updated  when the list is refactored. Only affected formulas
   are recomputed and the user can view elements on which the result depends.}
  {Collaborative Editing, Incremental Recomputation, Schema Change Control, End-User Debugging}
\extablecol{Hello World}{\cite{miller-2001-simult}}
  {Normalize the capitalization of two word messages.}
  {An operation to normalize the text in a list item can be recorded and applied to
   all list items or, alternatively, applied directly to all list items.}
  {Programming by Demonstration}
\extablecol{Traffic Accidents}{\cite{edwards-2022-copypaste}}
  {Compute statistics using two data sources.}
  {Formula to compute statistics can be reused with a different data source;
   error correction is propagated automatically to the copied version of the formula.}
  {Concrete Programming, Incremental Recomputation}
\hline
\end{tabular}

\caption{Formative examples used in Denicek design}
\label{fig:examples}
\end{figure*}
